---
title: "Live Session Unit 3 Assignment"
author: "Jasmine Coleman"
date: "January 30, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###1. GitHub Cloning
```{r GitClone}
# OrigDir <- getwd()
# newLocation <- "C:\\Users\\Jazzy\\OneDrive\\Documents\\SMU\\Spring '19\\Doing Data Science\\Unit 3"
# setwd(newLocation)
# getwd()
# dir.create("HW3")
# setwd("HW3")
# getwd()
# 
# gitRepo <- "https://github.com/caesar0301/awesome-public-datasets"
# git2r::clone(gitRepo,getwd())
```

  
###2. Data Summary  

####a.) This dataset follows the passengers aboard the Titanic, including their fees paid, rooms rented, and survivorship status.
```{r Read Titanic}
df <- read.csv("C:\\Users\\Jazzy\\OneDrive\\Documents\\SMU\\Spring '19\\Doing Data Science\\Unit 3\\HW3\\Datasets\\titanic.csv")
head(df)
```
  
####b.) Count and plot of females and males aboard the Titanic:
```{r Titanic Gender}
#install.packages('plyr')
library(plyr)

GenderCt <- count(df$Sex)
barplot(GenderCt$freq, names.arg = GenderCt$x, main = "Total Males and Females Aboard the Titanic", ylab = "Count", ylim = c(0,600))
```
  
####c.) Using one *apply* function to find the means of Age, Fare, and Survival.
```{r Titanic Apply}
MeanApply <-list(Age =na.omit(df$Age), Fare = df$Fare, Survival = df$Survived)
lapply(MeanApply, mean)
```

  
###3. Function Building
```{r Sleep Function}
sleep <- read.csv("C:\\Users\\Jazzy\\OneDrive\\Documents\\SMU\\Spring '19\\Doing Data Science\\Unit 3\\HW3\\sleep_data_01.csv")
head(sleep)

SleepFunction <- function(x)
{
  MedAge = median(na.omit(sleep$Age))
  minSleep = min(na.omit(sleep$Duration))
  maxSleep = max(na.omit(sleep$Duration))
  avgRSES = mean(na.omit(sleep$RSES))
  sdRSES = sd(na.omit(sleep$RSES))
  report <- data.frame(
    MedAge, avgRSES/5,sdRSES/5, maxSleep-minSleep
      )
  colnames(report) <- c("MedianAge", "SelfEsteem", "SE_SD", "DurationRange")
  round(report, 2)
}

SleepFunction()
```

  
###4. FiveThirtyEight Data  

####a.)	Installing the fivethirtyeight package
```{r 538Data Install}
#install.packages("fivethirtyeight")
library(fivethirtyeight)
```

####b.) Assigning the 22nd data set in the *listing of Data sets in package ‘fivethirtyeight,’* to object ‘df2.’
```{r 538Data Assign}
df2 <- college_recent_grads
```

####c.) Using a *more detailed list of the data sets* to identify the URL of a related news story.
```{r 538Data URL}
#vignette("fivethirtyeight", package = "fivethirtyeight")

#The news story related to this data can be found here:
#http://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/
```

####d.) Dimensions and column names of this dataframe.
```{r 538Data Columns}
dim(df2)
colnames(df2)
```

  
###Q5. Data Summary  

####a.) Column names and column count
```{r df2}
colnames(df2)
ncol(df2)
```

####b.) Recent Grad Major Category
```{r df2 Major Count}
major_count <- count(df2$major_category)
```

####c.) Barplot of Major Counts
```{r df2 Barplot}
par(las=2, mar=c(4, 11, 3, 1) + 0.1)
barplot(major_count$freq, names.arg = major_count$x, main = "Majors of Recent Grads", xlab = "Count", xlim = c(0,30),
        col = "blue", border = "pink", horiz = TRUE, cex.names = .7, cex.axis = .8)
```

####d.) Writing FiveThirtyEight data to a .csv file
```{r df2 .csv}
#write.csv(df2,file="fivethirtyeight_RecentCollegeGrad.csv", row.names = FALSE)
```


###Q6. Codebook
* Link to Jasmine Coleman's GitHub repository for MSDS DDS Spring 2019: https://github.com/jncolema/MSDS_DDS_Spring_2019
* Homework can be found in the "Homework" subdirectory.