---
title: "Live Session 5"
author: "Jasmine Coleman"
date: "February 12, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##GitHub Location:
https://github.com/jncolema/MSDS_DDS_Spring_2019/tree/master/Homework/Unit%205
Codebook can be found here as well.

Files and packages used in this assignment:
```{r download}
library(downloader)
library(dplyr)
# download("https://raw.githubusercontent.com/BivinSadler/MSDS-6306-Doing-Data-Science/master/Unit%205/yob2016.txt", destfile="yob2016.txt")
# download("https://raw.githubusercontent.com/BivinSadler/MSDS-6306-Doing-Data-Science/master/Unit%205/yob2015.txt", destfile="yob2015.txt")
# list.files()
```

##1. Data Munging  
###a. Importing yob2016.txt and Creating Dataframe
```{r import 2016}
df <- read.delim2("yob2016.txt", header = FALSE, sep = ";")
names(df) <- c("First_Name", "Gender", "Amount_of_Children")
head(df)
```

###b. df Structure
```{r df Structure}
head(df)
dim(df)
summary(df)
str(df)
```

###c. Finding Misspelled Entry
```{r Find Error}
which(grepl("yyy+", df$First_Name))
df[212,,]
```

###d. Removing Entry
```{r Remove Error}
y2016 <- df[-212,,]
```


##2. Data Merging  
###a. Importing yob2015.txt and Creating Dataframe
```{r Import 2015}
y2015 <- read.delim2("yob2015.txt", header = FALSE, sep = ",")
head(y2015)
dim(y2015)

names(y2015) <- c("First_Name", "Gender", "Amount_of_Children")
head(y2015)
```

###b. Displaying Last 10 Rows of Dataframe
```{r Last 2015}
tail(y2015,10)
```
####The last 10 rows are all for names stating with Z. What's interesting about them is that they were all only used 5 times and I've never heard of any of them.

###c. Merging y2016 and 72015
```{r Merge}
final <- merge(y2016, y2015, by = "First_Name")
head(final)
```

##3. Data Summary
###a. Creating Total Column
```{r Create Total}
final$Total <- final$Amount_of_Children.x+final$Amount_of_Children.y
head(final)
```

###b. Sorting by Total
```{r Sort Total}
head(final[order(-final$Total),],10)
```

###c. Omitting Boy Names
```{r Girls Only}
finalFemale <- filter(final,Gender.x=='F' & Gender.y=='F')
head(finalFemale)

Top10Female <- head(finalFemale[order(-finalFemale$Total),],10)
Top10Female$First_Name
```

###d. Writing Top 10 Girl Names to CSV File
```{r CSV}
# Top10Female$First_Name
# write.table(Top10Female$First_Name, file = "Top 10 Girl Names in 2015 and 2016.csv", sep = "," ,row.names=FALSE, col.names=FALSE)
```




