---
title: "Budweiser Competitive Analysis"
author: "Jasmine Coleman"
date: "February 28, 2019"
output:
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Purpose
####This code explores interesting findings on the US craft brewery market to help provide insight to Budweiser executives regarding future beer production.  

#Gathering the Data  
##Downloading the raw beer and brewery data  
```{r downloading}
# library(downloader)
# download("https://raw.githubusercontent.com/BivinSadler/MSDS-6306-Doing-Data-Science/master/Unit%207/Breweries.csv", destfile="Breweries.csv")
# download("https://raw.githubusercontent.com/BivinSadler/MSDS-6306-Doing-Data-Science/master/Unit%207/Beers.csv", destfile="Beers.csv")

```
  
##Importing the downloaded data
```{r importing}
Breweries <- read.csv2("Breweries.csv", header = TRUE, sep = ",")
dim(Breweries)
head(Breweries)
str(Breweries)
class(Breweries)

Beer <- read.csv2("Beers.csv", header = TRUE, sep = ",")
dim(Beer)
head(Beer)
str(Beer)
class(Beer)
```
  
#Analyzing the Data  
##How many breweries are in each state?
```{r BrewCount}
library(dplyr)
library(ggplot2)

#Counting occurrences for each state in dataset
BrewbyState <- Breweries %>% group_by(State) %>% count(State)

#Remove spaces from state variable
BrewbyState$State <- gsub(" ", "", BrewbyState$State, fixed = TRUE)

#Calculating % of total breweries in each state
BrewbyState$Percentage <- round((BrewbyState$n/sum(BrewbyState$n))*100,2)

#Putting results into a dataframe
BrewbyState <- data.frame(BrewbyState)
names(BrewbyState)[2] <- "Count"
str(BrewbyState)
summary(BrewbyState)

###Top 10 Brewery States###
TopBrewStates <- head(BrewbyState[order(-BrewbyState$Percentage),],10)

#Sorting table in descending order by percentage of total breweries
BrewbyState <- BrewbyState[order(-BrewbyState$Percentage),]
#Precentage of breweries located in 10 states with the most breweries
sum(head(BrewbyState$Percentage,10))

#exporting dataframe to csv file for Tableau heatmapping
#write.csv(BrewbyState, file = "BreweryCount.csv")
```

###This table displays how many breweries are in each state:
```{r Brew Table}
BrewbyState
```

###Plots visualizing brewery count by state:
```{r Brew Plots}
#Plots brewery count by state in descending order
ggplot(BrewbyState, aes(y=BrewbyState$Count , x=reorder(BrewbyState$State, BrewbyState$Count), fill=Count)) + geom_col() + coord_flip() +
  theme(plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), axis.title = element_text(size = 9, face = "bold"), axis.text.y = element_text(size=6),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line.x = element_line(colour = "black")) +
  labs(title = "Brewery Count by State", y = "Count", x = "State") +
  scale_fill_gradientn(colours = rainbow(3)) +
  scale_y_continuous(expand = c(0, 0))

#Plot brewery percentages by state
ggplot(BrewbyState, aes(y=BrewbyState$Percentage , x=reorder(BrewbyState$State, BrewbyState$Percentage), fill=Percentage)) + geom_col() + coord_flip() +
  theme(plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), axis.title = element_text(size = 9, face = "bold"), axis.text.y = element_text(size=6),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line.x = element_line(colour = "black")) +
  labs(title = "Percentage of Breweries in Each State", y = "Percentage", x = "State") +
  scale_fill_gradientn(colours = rainbow(3)) +
  scale_y_continuous(expand = c(0, 0))
```

##Merging Beer and Brewery Datasets
```{r Merging}
Beer_and_Brewery <-merge(Beer,Breweries, by.x = "Brewery_id", by.y = "Brew_ID", all.x = TRUE)
str(Beer_and_Brewery)
#Modifying variable types and updating column names
Beer_and_Brewery$State <- gsub(" ", "", Beer_and_Brewery$State, fixed = TRUE)
Beer_and_Brewery$ABV <- as.numeric(as.character(Beer_and_Brewery$ABV))
Beer_and_Brewery$Name.x <- as.character(Beer_and_Brewery$Name.x)
Beer_and_Brewery$Style <- as.character(Beer_and_Brewery$Style)
Beer_and_Brewery$Name.y <- as.character(Beer_and_Brewery$Name.y)
Beer_and_Brewery$City <- as.character(Beer_and_Brewery$City)
names(Beer_and_Brewery)[2] <- "Beer"
names(Beer_and_Brewery)[8] <- "Brewery"
str(Beer_and_Brewery)
```

###This is a sample of the merged dataset:
```{r Data Sample}
#Plotting First 6 and Last 6 Records

##First 6 rows of data
head(Beer_and_Brewery,6)

##Last 6 rows of data
tail(Beer_and_Brewery,6)

##Plotting First and Last 6 Rows of data
DataPreview <- data.frame(union(head(Beer_and_Brewery,6),tail(Beer_and_Brewery,6)))
DataPreview
```

###Plots visualizing beer count by state:
```{r Beer Count}
#Producing Table of Beers per State
BeerbyState <- Beer_and_Brewery %>% group_by(State) %>% summarize(no_rows = length(State))
BeerbyState$Percentage <- round((BeerbyState$no_rows/sum(BeerbyState$no_rows))*100,2)
names(BeerbyState)[2] <- "BeersBrewed" 
BeerbyState <-data.frame(BeerbyState)
BeerbyState

#Plotting beer production percentages by state
ggplot(BeerbyState, aes(y=BeerbyState$Percentage , x=reorder(BeerbyState$State, BeerbyState$Percentage), fill=Percentage)) + geom_col() + coord_flip() +
  theme(plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), axis.title = element_text(size = 9, face = "bold"), axis.text.y = element_text(size=6),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line.x = element_line(colour = "black")) +
  labs(title = "Percentage of Beers Brewed in Each State", y = "Percentage", x = "State") +
  scale_fill_gradientn(colours = rainbow(3)) +
  scale_y_continuous(expand = c(0, 0))

###Top 10 Beer States###
TopBeerStates<- head(BeerbyState[order(-BeerbyState$Percentage),],10)
TopBeerStateNames <- TopBeerStates[1:10,1]

#Sorting Beer Total by State Table
BeerbyState <- BeerbyState[order(-BeerbyState$Percentage),]
#Total percentage for top 10 brewery states
sum(head(BrewbyState$Percentage,10))
```
####Approximately 51% of all craft beer is produced in only 10 states.
```{r Top10 Plots}
#Plotting TopBrewStates
ggplot(TopBrewStates, aes(y=TopBrewStates$Count , x=reorder(TopBrewStates$State, TopBrewStates$Count), fill=-Count)) + geom_col() + coord_flip() +
  theme(plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), axis.title = element_text(size = 9, face = "bold"), axis.text.y = element_text(size=8, face = "bold"), axis.text.x = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line.x = element_blank(), axis.ticks.x = element_blank(), legend.position="none") +
  labs(title = "Top 10 Brewery States", y = "Total Breweries in Each State", x = "State") +
  scale_y_continuous(expand = c(0, 0),limits = c(0, 50), position = "right") +
  geom_text(aes(label = Count), hjust = -.5, size=3) #+ theme(legend.position="bottom")

#Plotting TopBeerStates
ggplot(TopBeerStates, aes(y=TopBeerStates$BeersBrewed , x=reorder(TopBeerStates$State, TopBeerStates$BeersBrewed), fill=-BeersBrewed)) + geom_col() + coord_flip() +
  theme(plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), axis.title = element_text(size = 9, face = "bold"), axis.text.y = element_text(size=8, face = "bold"), axis.text.x = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line.x = element_blank(), axis.ticks.x = element_blank(), legend.position="none") +
  labs(title = "Top 10 Beer Producing States", y = "Total Beers Produced", x = "State") + 
  scale_y_continuous(expand = c(0, 0),limits = c(0, 300), position = "right") +
  geom_text(aes(label = BeersBrewed), hjust = -.5, size=3) #+ theme(legend.position="bottom")
```

##Idenifying Any Missing Values
```{r Finding N/As}
#Counting N/As by column
apply(is.na(Beer_and_Brewery), 2, sum)
#Calculaing Percentage of N/As
NAs <- colSums(is.na(Beer_and_Brewery))
PercentNA <- (NAs/2410)*100
NAs
PercentNA
```
####There are 62 missing values for ABV (2.6%) and 1005 missing values for IBU (41.7%).

##Median Alcohol Content and Int'l Bitterness Units by State
```{r Median Values}
#Calculating Median Values
medianIBU <- aggregate(IBU~State,data=Beer_and_Brewery,median)
medianABV <- aggregate(as.numeric(as.character(Beer_and_Brewery$ABV))~State,data=Beer_and_Brewery,median)

#Merging both median vectors into one dataframe
medianstats <- merge(medianIBU, medianABV, by = "State", all = TRUE)
medianstats <- data.frame(medianstats)
names(medianstats)[3] <- "ABV"
dim(medianstats)
str(medianstats)
medianstats
```
####There is no IBU data available for craft beers produced in South Dakota.
```{r Median Barcharts}
#Bar Chart for Median IBU by State
ggplot(medianstats, aes(y=medianstats$IBU, x=reorder(medianstats$State, medianstats$IBU), fill=factor(IBU)))+
  geom_col(position="dodge", na.rm = FALSE) + coord_flip() +
  theme(plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), axis.title = element_text(size = 9, face = "bold"), axis.text.y = element_text(size=8, face = "bold"), axis.text.x = element_text(size=8, face = "bold"), axis.line.x = element_line(colour = "black"),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), legend.position="none") +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Median IBU by State", x = "State", y = "Median Values")

#Bar Chart for Median ABV by State
ggplot(medianstats,aes(y=medianstats$ABV, x=reorder(medianstats$State, medianstats$ABV), fill=factor(ABV)))+
  geom_col(position="dodge", na.rm = FALSE) + coord_flip() +
  theme(plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), axis.title = element_text(size = 9, face = "bold"), axis.text.y = element_text(size=8, face = "bold"), axis.text.x = element_text(size=8, face = "bold"), axis.line.x = element_line(colour = "black"),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), legend.position="none") +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Median ABV by State", x = "State", y = "Median Values")
```

####Plots of Top 10 Beer States' Median Values
```{r Top Median Plots}
#Subsetting data for top 10 states producing the most beer
TopBeerMedians <- subset(medianstats, State %in% TopBeerStateNames)

ABVPlot<- ggplot(TopBeerMedians,aes(y=TopBeerMedians$ABV, x=reorder(TopBeerMedians$State, TopBeerMedians$ABV), fill=factor(ABV)))+
  geom_col(position="dodge", na.rm = FALSE) + coord_flip() +
  theme(plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), axis.title = element_text(size = 9, face = "bold"), axis.text.y = element_text(size=8, face = "bold"), axis.text.x = element_blank(), axis.line.x = element_blank(), axis.ticks.x = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), legend.position="none") +
  scale_y_continuous(expand = c(0, 0),limits = c(0, .067), position = "right") +
  labs(title = "Median ABV for Top 10 Beer Producing States", x = "State", y = "Median Values") +
  geom_text(aes(label = ABV), hjust = -.5, size=3)

IBUPlot <- ggplot(TopBeerMedians, aes(y=TopBeerMedians$IBU, x=reorder(TopBeerMedians$State, TopBeerMedians$IBU), fill=factor(IBU)))+
  geom_col(position="dodge", na.rm = FALSE) + coord_flip() +
  theme(plot.title = element_text(hjust = 0.5, size = 11, face = "bold"), axis.title = element_text(size = 9, face = "bold"), axis.text.y = element_text(size=8, face = "bold"), axis.text.x = element_blank(), axis.line.x = element_blank(), axis.ticks.x = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), legend.position="none") +
  scale_y_continuous(expand = c(0, 0),limits = c(0, 45), position = "right") +
  labs(title = "Median IBU for Top 10 Beer Producing States", x = "State", y = "Median Values") +
  geom_text(aes(label = IBU), hjust = -.5, size=3)

ABVPlot
IBUPlot
```

###States Producing Beer with Maximum IBU and ABV Measurements
```{r Max Values}
#State with highest ABV and IBU values
maxIBUs <- aggregate(Beer_and_Brewery$IBU~State,data = Beer_and_Brewery,max)
maxABV <- aggregate(Beer_and_Brewery$ABV~State,data = Beer_and_Brewery,max)

head(maxIBUs[order(-maxIBUs$`Beer_and_Brewery$IBU`),],5)
head(maxABV[order(-maxABV$`Beer_and_Brewery$ABV`),],5)
```
####Oregon produces beer with the highest level of bitterness at 138 IBUs.
####Colorado produces beer with the highest alcohol content at 12.8% ABV.  
  
###ABV Summary Statistics
```{r ABV Stats}
summary(Beer_and_Brewery$ABV)
```
###IBU Summary Statistics
```{r IBU Stats}
summary(Beer_and_Brewery$IBU)
```

###Relationship Between Alcohol Content and Level of Bitterness
```{r Scatterplot}
#scatter plot to show the relationship between the bitterness of the beer and its alcoholic content
ggplot(Beer_and_Brewery, aes(x=ABV, y=IBU)) + geom_point(color="blue", size=1.5, alpha=0.3, pch=16) +
  theme_bw(base_size=16) + geom_smooth(method = "lm", se = T, color="red")

#Correlation Test
cor.test(Beer_and_Brewery$ABV,Beer_and_Brewery$IBU, alternative = "two.sided", method = "pearson", conf.level = 0.95)
#Pearson's r
cor(Beer_and_Brewery$ABV,Beer_and_Brewery$IBU, use="complete.obs", method = "pearson")
#Covariance
cov(Beer_and_Brewery$ABV,Beer_and_Brewery$IBU, use="complete.obs", method = "pearson")
```
####There is a strong positive linear relationship between ABV and IBU (r = 0.67, *p* = 2.2e-16 from Pearson's r) . As the alcohol content of beer increases, the level of bitterness in the beer also tends to increase. 

###Summary
####Potential Target Markets
- 10 States housing ~50% of all craft breweries  
- 10 States producing ~50% of all craft beers  

####Ideal Beer Measurements  
- ABV: 5.2% - 6.2%  
- IBV: 19 - 42  
- ABV and IBU are positively correlated  

####Future Research  
- Obtain missing IBU values or understand why so many are missing  

###Links
- The YouTube presentation for this analysis can be found here: https://youtu.be/BKGMBUrEUE4  
- The GitHub repo for this analysis can be found here: https://github.com/jncolema/MSDS_DDS_Spring_2019/tree/master/Project  